### 2024-11-17 scripts/05-exploratory_data_analysis.qmd
Upload draft for scripts
all seven scripts to clean data, analysis data, simulated data, explore data and create model

---
title: "05-exploratory_data_analysis"
author: "Shaotong LI"
format: pdf
editor: visual
---

# Preamble

#### Purpose: Explore the analysis data

#### Author: Shaotong Li

#### Date: 11 November 2024

#### Contact: maxst.li\@mail.utoronto.ca

#### License: MIT

#### Pre-requisites: analysis data generated in 02-analysis_data

#### Any other information needed? None

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#### Workspace setup ####

#Install necessary packages
install.packages("tidyverse")
install.packages("arrow")
install.packages("here")
install.packages("ggplot2")
install.packages("patchwork")
install.packages("tidyverse")
install.packages("sf")
install.packages("rnaturalearth")
install.packages("rnaturalearthdata")
install.packages("ggthemes")
install.packages("dplyr")
install.packages("knitr")

# Load libraries
library(tidyverse)
library(arrow)
library(here)
library(ggplot2)
library(patchwork)
library(tidyverse)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggthemes)
library(dplyr)
library(knitr)

# Load cleaned datasets
sakura_historical_cleaned <- read_csv(here::here("data/03-cleaned_data/sakura-historical-cleaned.csv"))
sakura_modern_cleaned <- read_csv(here::here("data/03-cleaned_data/sakura-modern-cleaned.csv"))
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#### Plot Distribution of Flower Dates and DOY by Month and Day (Continuous Timeline) ####

# Extract month and day from flower_date
flower_date_distribution <- sakura_historical_cleaned %>%
  mutate(
    month_num = month(flower_date),  # Extract month as a number
    day_of_year = yday(flower_date)  # Create a continuous day-of-year timeline
  )

# Group the data to calculate flower_date counts and average flower_doy
flower_date_summary <- flower_date_distribution %>%
  group_by(day_of_year) %>%
  summarise(
    flower_date_count = n(),
    flower_doy_mean = mean(flower_doy, na.rm = TRUE),
    .groups = "drop"
  )

# Plot the data with two lines
flower_date_plot <- ggplot(flower_date_summary, aes(x = day_of_year)) +
  geom_line(aes(y = flower_date_count, color = "Flower Date Count"), size = 1) +  # Line for flower_date count
  geom_line(aes(y = flower_doy_mean, color = "Average Flower DOY"), size = 1) +  # Line for flower_doy
  scale_color_manual(
    values = c("Flower Date Count" = "steelblue", "Average Flower DOY" = "darkorange"),
    name = "Legend"
  ) +
  labs(
    title = "Flower Date Distribution and DOY",
    subtitle = "Flower date count and average flower DOY across days of the year",
    x = "Day of the Year",
    y = "Count / DOY"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 12),
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12),
    legend.position = "right"
  )

# Display the plot
print(flower_date_plot)

```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#### Prepare Data ####
# sakura_historical is already loaded
scatter_data <- sakura_historical_cleaned %>%
  filter(!is.na(temp), !is.na(flower_doy))  # Ensure no missing values

#### Create Scatter Plot ####
scatter_plot_no_line <- ggplot(scatter_data, aes(x = temp, y = flower_doy)) +
  geom_point(aes(color = year), alpha = 0.6, size = 2) +  # Scatter points colored by year
  scale_color_viridis_c(option = "C", name = "Year") +  # Color scale for year
  labs(
    title = "Relationship Between Temperature and Flowering Day",
    subtitle = "Color represents temporal progression",
    x = "Temperature (°C)",
    y = "Flowering Day (DOY)"
  ) +
  theme_minimal() +
  theme(
    legend.position = "right",
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 12),
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12)
  )

# Display the plot
print(scatter_plot_no_line)

```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#### Prepare Data ####
# sakura_historical_cleaned is already loaded
line_data <- sakura_historical_cleaned %>%
  filter(!is.na(temp), !is.na(year))  # Ensure no missing values

#### Create Line Plot ####
line_plot <- ggplot(line_data, aes(x = year, y = temp)) +
  geom_line(color = "steelblue", size = 1) +  # Line graph with color and size
  labs(
    title = "Temperature Trend Over Years",
    subtitle = "Yearly average temperature",
    x = "Year",
    y = "Temperature (°C)"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 12),
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12)
  )

# Display the plot
print(line_plot)

```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#### Improved Plot Latitude and Longitude Distributions ####

# Plot latitude distribution with bins
latitude_plot <- ggplot(sakura_modern_cleaned, aes(x = latitude)) +
  geom_histogram(
    bins = 20,                    # Adjust the number of bins for better grouping
    fill = "steelblue", 
    alpha = 0.8
  ) +
  labs(
    title = "Latitude Distribution",
    x = "Latitude",
    y = "Count"
  ) +
  theme_minimal()

# Plot longitude distribution with bins
longitude_plot <- ggplot(sakura_modern_cleaned, aes(x = longitude)) +
  geom_histogram(
    bins = 20,                    # Adjust the number of bins for better grouping
    fill = "darkorange", 
    alpha = 0.8
  ) +
  labs(
    title = "Longitude Distribution",
    x = "Longitude",
    y = "Count"
  ) +
  theme_minimal()

# Display the plots side by side
latitude_plot + longitude_plot

```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#### Normalize and Prepare Data ####
# Longitude data
longitude_temp_data <- sakura_modern_cleaned %>%
  filter(!is.na(longitude), !is.na(mean_temp_month)) %>%
  group_by(longitude) %>%
  summarise(mean_temp = mean(mean_temp_month, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(type = "Longitude")  # Mark this as Longitude data

# Latitude data
latitude_temp_data <- sakura_modern_cleaned %>%
  filter(!is.na(latitude), !is.na(mean_temp_month)) %>%
  group_by(latitude) %>%
  summarise(mean_temp = mean(mean_temp_month, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(type = "Latitude")  # Mark this as Latitude data

# Rescale latitude for alignment (latitude scaled to match longitude range)
latitude_temp_data <- latitude_temp_data %>%
  mutate(latitude_rescaled = scales::rescale(latitude, to = range(longitude_temp_data$longitude)))

#### Combine Data for Plotting ####
# Prepare data for plotting
longitude_temp_data <- longitude_temp_data %>%
  rename(variable = longitude)

latitude_temp_data <- latitude_temp_data %>%
  rename(variable = latitude_rescaled)

combined_temp_data <- bind_rows(
  longitude_temp_data,
  latitude_temp_data
)

#### Create Plot ####
ggplot() +
  # Longitude line
  geom_line(
    data = longitude_temp_data,
    aes(x = variable, y = mean_temp),
    color = "steelblue", size = 1
  ) +
  # Latitude line (rescaled)
  geom_line(
    data = latitude_temp_data,
    aes(x = variable, y = mean_temp),
    color = "darkorange", size = 1
  ) +
  # Add axis labels and legend
  labs(
    title = "Mean Temperature by Longitude and Latitude",
    subtitle = "Relationship between Longitude/Latitude and mean temperature",
    x = "Longitude / Rescaled Latitude",
    y = "Mean Temperature (°C)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 12),
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12)
  ) +
  # Add a secondary x-axis for latitude
  scale_x_continuous(
    name = "Longitude",
    sec.axis = sec_axis(
      ~scales::rescale(., to = range(latitude_temp_data$latitude)),
      name = "Latitude"
    )
  )
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Fit a linear model: mean_temp_month ~ day + latitude + month
model3 <- lm(mean_temp_month ~ latitude + longitude, data = sakura_modern_cleaned)

# Display summary of the combined model
cat("\nCombined Model: Linear model for Mean Temperature ~ Day + Latitude + Month\n")
print(summary(model3))

```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#### Plot Monthly Distribution of Flower Date ####

# Extract the month from flower_date and create a summary table
flower_month_distribution <- sakura_modern_cleaned %>%
  mutate(month = month(flower_date, label = TRUE, abbr = TRUE)) %>%  # Extract abbreviated month names
  count(month)  # Count the occurrences for each month

# Plot the monthly distribution
flower_month_plot <- ggplot(flower_month_distribution, aes(x = month, y = n)) +
  geom_bar(stat = "identity", fill = "steelblue", alpha = 0.8) +  # Use steelblue for the bars
  labs(
    title = "Monthly Distribution of Flower Date",
    x = "Month",
    y = "Count"
  ) +
  theme_minimal()

# Display the plot
flower_month_plot
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#### Plot Distribution of Flower Dates by Month and Day ####

# Extract month and day from flower_date
flower_date_distribution <- sakura_modern_cleaned %>%
  mutate(month = month(flower_date, label = TRUE, abbr = TRUE),  # Extract abbreviated month names
         day = day(flower_date))  # Extract day of the month

# Create a summarized dataset for the count of flower dates by month and day
flower_date_summary <- flower_date_distribution %>%
  group_by(month, day) %>%
  summarise(count = n(), .groups = "drop")

# Use a broader color palette for all months
color_palette <- scales::hue_pal()(length(unique(flower_date_summary$month)))

# Plot the line graph
flower_date_line_plot <- ggplot(flower_date_summary, aes(x = day, y = count, color = month, group = month)) +
  geom_line(size = 1) +  # Draw smooth lines for each month
  scale_color_manual(values = color_palette) +  # Use a color palette for all months
  labs(
    title = "Flower Date Distribution by Month and Day",
    x = "Day of the Month",
    y = "Count",
    color = "Month"
  ) +
  theme_minimal() +
  theme(
    legend.position = "right"
  )

# Display the plot
flower_date_line_plot
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#### Load and Prepare Data ####
# Calculate mean flower_doy for each location
sakura_summary <- sakura_modern_cleaned %>%
  group_by(latitude, longitude) %>%
  summarise(mean_flower_doy = mean(flower_doy, na.rm = TRUE)) %>%
  ungroup()

#### Load Japan's Basemap ####
japan_map <- ne_countries(scale = "medium", country = "Japan", returnclass = "sf")

#### Create the Map ####
gis_plot <- ggplot(data = japan_map) +
  geom_sf(fill = "white", color = "black") +  # Base map of Japan
  geom_point(
    data = sakura_summary,
    aes(x = longitude, y = latitude, color = mean_flower_doy),
    size = 3, alpha = 0.7
  ) +
  scale_color_viridis_c(
    option = "D",  # Viridis color scale for better visualization
    name = "Mean Flower DOY"
  ) +
  theme_minimal() +
  labs(
    title = "Average Flowering Dates Across Japan",
    subtitle = "Darker points indicate later flowering dates",
    x = "Longitude",
    y = "Latitude"
  ) +
  theme(legend.position = "right")

# Display the plot
print(gis_plot)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#### Load and Prepare Data ####
# Calculate mean temperature for each location
temperature_summary <- sakura_modern_cleaned %>%
  group_by(latitude, longitude) %>%
  summarise(mean_temp_month = mean(mean_temp_month, na.rm = TRUE)) %>%
  ungroup()

#### Load Japan's Basemap ####
japan_map <- ne_countries(scale = "medium", country = "Japan", returnclass = "sf")

#### Create the Map for Temperature ####
gis_plot_temp <- ggplot(data = japan_map) +
  geom_sf(fill = "white", color = "black") +  # Base map of Japan
  geom_point(
    data = temperature_summary,
    aes(x = longitude, y = latitude, color = mean_temp_month),
    size = 3, alpha = 0.7
  ) +
  scale_color_gradientn(
    colors = c("#2c7bb6", "#abd9e9", "#ffffbf", "#fdae61", "#d7191c"),  # Blue to red gradient
    name = "Mean Temperature (°C)"
  ) +
  theme_minimal() +
  labs(
    title = "Average Monthly Temperatures Across Japan",
    subtitle = "Warmer colors indicate higher temperatures",
    x = "Longitude",
    y = "Latitude"
  ) +
  theme(legend.position = "right")

# Display the plot
print(gis_plot_temp)

```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#### Plot Distribution of Temperature by Date ####

# Extract date and temperature from the dataset
temperature_distribution <- sakura_modern_cleaned %>%
  select(flower_date, mean_temp_month) %>%
  mutate(
    day_of_year = yday(flower_date),  # Convert date to day of the year
    month = month(flower_date, label = TRUE, abbr = TRUE)  # Extract month for potential color coding
  )

# Create the plot
temperature_plot <- ggplot(temperature_distribution, aes(x = day_of_year, y = mean_temp_month)) +
  geom_point(alpha = 0.5, color = "steelblue") +  # Scatter plot of temperature
  geom_smooth(method = "loess", color = "darkorange", size = 1, se = FALSE) +  # Smoothed line for trend
  labs(
    title = "Temperature Distribution by Date",
    subtitle = "Scatter plot with temperature trend over the year",
    x = "Day of the Year",
    y = "Temperature (°C)"
  ) +
  scale_x_continuous(
    breaks = c(1, 32, 60, 91, 121, 152),  # Manually specify breaks for Jan, Feb, Mar, Apr, May
    labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun")  # Correct labels
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 12),
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12)
  )

# Display the plot
print(temperature_plot)

```