### 2024-11-18 paper/Paper.qmd
Update Paper and reference (second draft)
This is the second draft of this paper. References are now complete. Paper contents are now all complete. Further change will be uploaded as third draft.

---
title: "Predicting Sakura Florescence in Japan with a Dual-Model Approach"
subtitle: "Estimating Temperature and Forecasting Bloom Duration for Cherry Blossoms"
author: 
  - Shaotong (Max) Li
thanks: "Code and data are available at: [Sakura_Florescence_Prediction](https://github.com/ShaotongLi-Max/Sakura_Florescence_Prediction.git)."
date: 2024/11/17
date-format: long
abstract: "This paper presents a dual-model approach to predict the duration of cherry blossom blooms in Japan: The first model predicts bloom duration using temperature, while the second model estimates temperature based on latitude, month, and date. By analyzing historical and modern data, we provide insights into how latitude and seasonal timing affect temperature and bloom duration. Our findings indicate that this dual-model approach enhances our understanding of climatic effects on cherry blossom florescence, highlighting potential impacts of climate change on one of Japan's most cherished natural events."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---
```{r}
#| include: false
#| warning: false
#| message: false
#| echo: false

#### Workspace setup ####
install.packages("tidyverse")
install.packages("arrow")
install.packages("here")
install.packages("ggplot2")
install.packages("patchwork")
install.packages("tidyverse")
install.packages("sf")
install.packages("rnaturalearth")
install.packages("rnaturalearthdata")
install.packages("ggthemes")
install.packages("dplyr")
install.packages("knitr")
install.packages("kableExtra")

# Load libraries
library(tidyverse)
library(arrow)
library(here)
library(ggplot2)
library(patchwork)
library(tidyverse)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggthemes)
library(dplyr)
library(knitr)
library(kableExtra)

# Load cleaned datasets
sakura_historical_cleaned <- read_csv(here::here("data/03-cleaned_data/sakura-historical-cleaned.csv"))
sakura_modern_cleaned <- read_csv(here::here("data/03-cleaned_data/sakura-modern-cleaned.csv"))
# Load temp_test_data for Model 1
temp_test_data <- read_parquet(here::here("data/02-analysis_data/temp_test_data.parquet"))
# Load latitude&time_test_data for Model 2
latitude_time_test_data <- read_parquet(here::here("data/02-analysis_data/latitude&time_test_data.parquet"))
# Load Model 1
model1 <- readRDS(here::here("models/model1.rds"))
# Load Model 2
model2 <- readRDS(here::here("models/model2.rds"))
```

# Introduction

Cherry blossoms, or sakura, are a powerful symbol of Japan, representing both the fleeting beauty of nature and a cherished cultural heritage [@Wikipedia2024]. Each spring, the blooming of cherry blossoms is celebrated across the country, marking a time of renewal and festivity [@Wikipedia2024]. However, the timing and duration of these blooms are highly sensitive to climatic conditions, which makes predicting sakura florescence increasingly important in the face of climate change [@Ocko2024]. Understanding how environmental factors, such as temperature, latitude, and seasonal timing, influence bloom duration can help forecast these events and provide insight into broader ecological changes [@Ocko2024].

This paper aims to address the challenge of predicting cherry blossom bloom duration by using a dual-model approach. The first model predicts the duration of cherry blossom blooms using temperature. The second model estimates temperature based on environmental variables such as latitude, month, and date, providing a detailed understanding of local climatic conditions. By combining these two models, we seek to fill the gap in understanding how temperature fluctuations driven by geographic and temporal factors affect sakura florescence.

The estimand in this study is the predicted duration of cherry blossom (sakura florescence). Specifically, Model 1 predicts the duration of the bloom using temperature, while Model 2 estimates temperature as a function of latitude, month, and date. This dual-model structure allows us to precisely quantify the effect of temperature as a mediating factor influenced by geographical and temporal variables on bloom duration. This approach allows us to quantify the relationship between environmental conditions and the duration of cherry blossom blooms.

We utilized both historical and modern phenological data to ensure the robustness of our predictions. Our analysis highlights how latitude and seasonal factors contribute to temperature variation, which in turn affects the timing and duration of the bloom. The results demonstrate the potential for these models to improve the accuracy of bloom forecasts, particularly in the context of changing climatic patterns. This research contributes valuable knowledge to the fields of phenology and climate science, and provides practical implications for tourism, cultural events, and ecological conservation in Japan.

The remainder of this paper is structured as follows: @sec-data discusses the data sources and preprocessing methods. @sec-model details the dual-model approach, including the temperature estimation and bloom prediction models. @sec-result presents the results, followed by a discussion in @sec-discussion on the implications of our findings. Finally, @sec-appendix concludes with insights into future research directions and the broader impact of climate change on cherry blossoms.

# Data {#sec-data}

In this project, we used R[@citeR] and several R packages for data processing, analysis, and visualization. Specifically, tidyverse [@tidyverse], arrow[@arrow], here[@here], ggplot2[@ggplot2], patchwork[@patchwork], sf[@sf], rnaturalearth[@rnaturalearth], and rnaturalearthdata[@rnaturalearthdata] were used for data processing, geospatial analysis, and visualization. ggthemes[@ggthemes] was utilized to apply thematic elements to plots, while dplyr[@dplyr] was key for data manipulation tasks. For dynamic report generation, knitr[@knitr] and kableExtra[@kableExtra] were used, providing enhanced formatting for outputs. Together, these packages enabled efficient data cleaning, analysis, and visualization throughout the study.

## Overview

The data used in this study includes three primary datasets: historical cherry blossom bloom records, modern bloom records, and temperature data. The historical dataset  contains records of cherry blossom bloom dates across multiple decades for the Kyoto region, providing a long-term view of bloom trends in this specific area of Japan. The modern dataset  contains recent bloom dates from various regions across Japan, reflecting current climatic conditions and bloom dynamics. Lastly, the temperature dataset  includes temperature readings for various locations in Japan, which are essential for estimating temperature-related effects on bloom duration.

To ensure comparability, only bloom records that met our defined geographic and seasonal criteria were retained. This filtering process ensured that only reliable and relevant data points were used in the final analysis, thus maintaining the accuracy and robustness of our predictive models. Finally, the cleaned datasets were merged based on common identifiers such as location and time to create a comprehensive dataset for modeling sakura florescence.

The final combined dataset, after merging and filtering, consisted of 10,456 rows with multiple key attributes, including latitude, month, date, and temperature, serving as predictors for bloom duration. By integrating both historical and modern data, our analysis captures both long-term trends and recent shifts in bloom timing, providing insights into the effects of climate variability and change on cherry blossom phenology.

## Measurement

The transformation of real-world phenomena into dataset entries involved detailed documentation of cherry blossom bloom dynamics. Historical sakura data was derived from records such as diary entries and observations from meteorological agencies, specifically focusing on the Kyoto region. Bloom dates were reconstructed based on these historical accounts dating back to 812 CE, supplemented by temperature data that provided context for climatic influences over centuries.

Modern cherry blossom data was collected by the Japan Meteorological Agency (JMA), including flowering dates and meteorological information across Japan since 1953. This dataset encompasses latitude, longitude, bloom dates, and temperature data, offering a comprehensive view of bloom dynamics under current climatic conditions. Temperature readings were matched with bloom records using geographic and temporal identifiers to ensure consistent analysis.

Data quality checks included removing incomplete records, standardizing date formats, and verifying bloom dates against multiple sources. This ensured that the final dataset was robust and suitable for modeling, enabling reliable predictions of bloom duration based on environmental factors.

## Outcome Variables

### Florescence

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.pos="H"}
#| label: fig-florescence
#| fig-cap: "Sakura Florescence: Historical vs Modern Data"

#### Combine and Prepare Data ####
# Prepare sakura_historical_cleaned data
historical_flower_data <- sakura_historical_cleaned %>%
  select(flower_doy) %>%
  filter(!is.na(flower_doy)) %>%  # Remove missing values
  mutate(dataset = "Historical")  # Add label for distinction

# Prepare sakura_modern_cleaned data
modern_flower_data <- sakura_modern_cleaned %>%
  select(flower_doy) %>%
  filter(!is.na(flower_doy)) %>%  # Remove missing values
  mutate(dataset = "Modern")  # Add label for distinction

# Combine the datasets
combined_flower_data <- bind_rows(historical_flower_data, modern_flower_data)

#### Plot Flowering Day Distribution ####
flower_doy_distribution_plot <- combined_flower_data %>%
  ggplot(aes(x = flower_doy, fill = dataset, color = dataset)) +
  geom_density(alpha = 0.3, adjust = 1.5) +  # Use density plot for smoother curves
  labs(
    x = "Flowering Day (DOY)",
    y = "Density",
    fill = "Dataset",
    color = "Dataset"
  ) +
  scale_fill_manual(values = c("Historical" = "steelblue", "Modern" = "darkorange")) +
  scale_color_manual(values = c("Historical" = "steelblue", "Modern" = "darkorange")) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12),
    legend.position = "right"
  )

# Display the plot
print(flower_doy_distribution_plot)

```
The variable 'florescence' represents the duration of cherry blossom blooming, defined as the number of days during which cherry blossoms are in full bloom. In the historical dataset, which focuses on the Kyoto region, florescence data shows a narrower distribution, indicating relatively stable bloom timings over centuries. In contrast, the modern dataset includes bloom data from multiple regions across Japan, resulting in a wider distribution that reflects greater variability due to regional climatic differences.

@fig-florescence illustrates the density distribution of the day of year (DOY) for full bloom, comparing historical and modern datasets. The historical data (Kyoto) exhibits a peak around a consistent time frame, whereas the modern data shows more dispersed bloom times, likely driven by regional climatic variations and recent temperature changes.

As the outcome variable in our modeling approach, florescence is crucial for understanding bloom dynamics. It allows us to assess the impact of temperature and other environmental factors on the timing and length of cherry blossom blooms, providing insights into the effects of climate variability.

The distribution of florescence in the historical data shows relatively stable bloom timings, with less variability compared to the modern dataset. This stability is largely due to the consistent climatic conditions in the Kyoto region over centuries. In contrast, the modern data presents more pronounced regional variations in bloom duration, influenced by diverse environmental conditions across Japan. As the outcome variable in our modeling approach, florescence serves as the key measure of bloom dynamics, enabling us to evaluate how temperature and other environmental factors affect the timing and duration of cherry blossom blooms in both historical and modern contexts.

## Predictor Variables

### Temperature

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.pos="H"}
#| label: fig-temperature-1
#| fig-cap: "Temperature: Historical vs Modern Data"

#### Combine and Prepare Data ####
# Prepare sakura_historical_cleaned data
historical_temp_data <- sakura_historical_cleaned %>%
  select(temp) %>%
  filter(!is.na(temp)) %>%  # Remove missing values
  mutate(dataset = "Historical", variable = "temp")  # Add labels for distinction

# Prepare sakura_modern_cleaned data
modern_temp_data <- sakura_modern_cleaned %>%
  select(mean_temp_month) %>%
  filter(!is.na(mean_temp_month)) %>%  # Remove missing values
  rename(temp = mean_temp_month) %>%  # Rename column to match for combining
  mutate(dataset = "Modern", variable = "mean_temp_month")  # Add labels for distinction

# Combine the datasets
combined_temp_data <- bind_rows(historical_temp_data, modern_temp_data)

#### Plot Temperature Distribution ####
temp_distribution_plot <- combined_temp_data %>%
  ggplot(aes(x = temp, fill = dataset, color = dataset)) +
  geom_density(alpha = 0.3, adjust = 1.5) +  # Use density plot for smoother curves
  labs(
    x = "Temperature (°C)",
    y = "Density",
    fill = "Dataset",
    color = "Dataset"
  ) +
  scale_fill_manual(values = c("Historical" = "steelblue", "Modern" = "darkorange")) +
  scale_color_manual(values = c("Historical" = "steelblue", "Modern" = "darkorange")) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12),
    legend.position = "right"
  )

# Display the plot
print(temp_distribution_plot)

```

Temperature plays a crucial role in determining the timing and duration of cherry blossom blooms. @fig-temperature-1 shows the density distribution of temperatures in historical and modern datasets, highlighting significant differences between the two periods. The historical dataset, primarily focused on the Kyoto region, shows a relatively narrow temperature range centered around moderate values, while the modern dataset presents a broader range with lower temperatures being more prevalent. This difference in temperature distributions likely reflects the increased regional variability and recent climatic changes observed in Japan.

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.pos="H"}
#| label: fig-temperature-2
#| fig-cap: "Temperature VS Sakura Florescence"

#### Prepare Data ####
# sakura_historical is already loaded
scatter_data <- sakura_historical_cleaned %>%
  filter(!is.na(temp), !is.na(flower_doy))  # Ensure no missing values

#### Create Scatter Plot ####
scatter_plot_no_line <- ggplot(scatter_data, aes(x = temp, y = flower_doy)) +
  geom_point(aes(color = year), alpha = 0.6, size = 2) +  # Scatter points colored by year
  scale_color_viridis_c(option = "C", name = "Year") +  # Color scale for year
  labs(
    subtitle = "Color represents temporal progression",
    x = "Temperature (°C)",
    y = "Flowering Day (DOY)"
  ) +
  theme_minimal() +
  theme(
    legend.position = "right",
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 12),
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12)
  )

# Display the plot
print(scatter_plot_no_line)

```

The strong linear relationship between temperature and the day of year (DOY) for full bloom is evident in @fig-temperature-2. As temperatures increase, the DOY for blooming decreases, indicating that higher temperatures lead to earlier blooming. This relationship is consistent across both historical and modern datasets, underscoring the critical impact of temperature on bloom timing. The temporal progression in @fig-temperature-2, represented by the color gradient, illustrates how this relationship has persisted over time, even as overall climatic conditions have changed.

In our modeling approach, temperature serves as a key predictor for florescence, allowing us to quantify how fluctuations in temperature directly affect the timing and duration of cherry blossom blooms. By examining the historical and modern temperature distributions, we can better understand the impacts of climate variability and identify trends that are crucial for predicting future bloom behavior.

### Latitude

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.pos="H"}
#| label: fig-latitude-1
#| fig-cap: "Latitude Distribution"

#### Improved Plot Latitude and Longitude Distributions ####

# Plot latitude distribution with bins
latitude_plot <- ggplot(sakura_modern_cleaned, aes(x = latitude)) +
  geom_histogram(
    bins = 20,                    # Adjust the number of bins for better grouping
    fill = "steelblue", 
    alpha = 0.8
  ) +
  labs(
    x = "Latitude",
    y = "Count"
  ) +
  theme_minimal()

latitude_plot

```

@fig-latitude-1 shows the distribution of latitudes for the locations included in the dataset. The majority of observations are concentrated between latitudes 34° and 38° N, which represents the regions most commonly associated with cherry blossom observations in Japan. There is a smaller number of observations at lower and higher latitudes, reflecting the geographic range of sakura coverage in the country.

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.pos="H"}
#| label: fig-latitude-2
#| fig-cap: "Latitude VS Temperature"

#### Prepare Latitude Data ####
latitude_temp_data <- sakura_modern_cleaned %>%
  filter(!is.na(latitude), !is.na(mean_temp_month)) %>%
  group_by(latitude) %>%
  summarise(mean_temp = mean(mean_temp_month, na.rm = TRUE)) %>%
  ungroup()

#### Fit a Linear Model ####
linear_model <- lm(mean_temp ~ latitude, data = latitude_temp_data)

#### Create Plot ####
latitude_temp_plot <- ggplot(latitude_temp_data, aes(x = latitude, y = mean_temp)) +
  geom_line(color = "steelblue", size = 1) +  # Line graph with custom color and size
  geom_abline(
    slope = coef(linear_model)["latitude"],
    intercept = coef(linear_model)["(Intercept)"],
    color = "darkorange",
    linetype = "dashed",
    size = 1
  ) +  # Add a dashed line for the linear relationship
  labs(
    subtitle = "Relationship with Linear Trend",
    x = "Latitude",
    y = "Mean Temperature (°C)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 12),
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12)
  )

# Display the plot
print(latitude_temp_plot)

```

@fig-latitude-2 illustrates the relationship between latitude and mean temperature, showing a clear positive trend. As latitude increases, so does the average temperature. The dashed orange line represents the linear trend, indicating a consistent increase in temperature as we move towards the northern regions of Japan. This relationship is essential for understanding how geographic position influences local climatic conditions, which in turn affects bloom timing.

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.pos="H"}
#| label: fig-latitude-3
#| fig-cap: "Average Temperatures Across Japan"

#### Load and Prepare Data ####
# Calculate mean temperature for each location
temperature_summary <- sakura_modern_cleaned %>%
  group_by(latitude, longitude) %>%
  summarise(mean_temp_month = mean(mean_temp_month, na.rm = TRUE)) %>%
  ungroup()

#### Load Japan's Basemap ####
japan_map <- ne_countries(scale = "medium", country = "Japan", returnclass = "sf")

#### Create the Map for Temperature ####
gis_plot_temp <- ggplot(data = japan_map) +
  geom_sf(fill = "white", color = "black") +  # Base map of Japan
  geom_point(
    data = temperature_summary,
    aes(x = longitude, y = latitude, color = mean_temp_month),
    size = 3, alpha = 0.7
  ) +
  scale_color_gradientn(
    colors = c("#2c7bb6", "#abd9e9", "#ffffbf", "#fdae61", "#d7191c"),  # Blue to red gradient
    name = "Mean Temperature (°C)"
  ) +
  theme_minimal() +
  labs(
    subtitle = "Warmer colors indicate higher temperatures",
    x = "Longitude",
    y = "Latitude"
  ) +
  theme(legend.position = "right")

# Display the plot
print(gis_plot_temp)

```

@fig-latitude-3 presents a geographic distribution of mean temperatures across Japan, with warmer colors indicating higher temperatures. This visualization highlights the spatial temperature variability across the country, where northern regions tend to have higher average temperatures compared to southern regions. The color gradient helps to illustrate how temperature conditions vary as a function of both latitude and longitude.

In summary, latitude has a significant influence on temperature, which directly impacts the timing of cherry blossom blooms. The relationship between latitude and temperature is well-captured by Figures 4, 5, and 6, which collectively demonstrate that regions at higher latitudes tend to experience higher average temperatures, contributing to earlier bloom periods. These insights are crucial for understanding the geographic factors that affect sakura florescence and for improving the accuracy of predictive models.

### Time of Blossom

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.pos="H"}
#| label: fig-flowerday-1
#| fig-cap: "Monthly Distribution of Flower Date"

#### Plot Monthly Distribution of Flower Date ####

# Extract the month from flower_date and create a summary table
flower_month_distribution <- sakura_modern_cleaned %>%
  mutate(month = month(flower_date, label = TRUE, abbr = TRUE)) %>%  # Extract abbreviated month names
  count(month)  # Count the occurrences for each month

# Plot the monthly distribution
flower_month_plot <- ggplot(flower_month_distribution, aes(x = month, y = n)) +
  geom_bar(stat = "identity", fill = "steelblue", alpha = 0.8) +  # Use steelblue for the bars
  labs(
    x = "Month",
    y = "Count"
  ) +
  theme_minimal()

# Display the plot
flower_month_plot
```

@fig-flowerday-1 shows the monthly distribution of flower dates, with the majority of cherry blossoms blooming in March and April. This aligns with the typical cherry blossom season in Japan, which occurs in early spring. There are fewer occurrences of blooms in January, February, and May, which indicates that bloom timings outside of this window are uncommon, reflecting the strong seasonal pattern of sakura flowering.

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.pos="H"}
#| label: fig-flowerday-2
#| fig-cap: "Flower Date Distribution by Month and Day"

#### Plot Distribution of Flower Dates by Month and Day ####

# Extract month and day from flower_date
flower_date_distribution <- sakura_modern_cleaned %>%
  mutate(month = month(flower_date, label = TRUE, abbr = TRUE),  # Extract abbreviated month names
         day = day(flower_date))  # Extract day of the month

# Create a summarized dataset for the count of flower dates by month and day
flower_date_summary <- flower_date_distribution %>%
  group_by(month, day) %>%
  summarise(count = n(), .groups = "drop")

# Use a broader color palette for all months
color_palette <- scales::hue_pal()(length(unique(flower_date_summary$month)))

# Plot the line graph
flower_date_line_plot <- ggplot(flower_date_summary, aes(x = day, y = count, color = month, group = month)) +
  geom_line(size = 1) +  # Draw smooth lines for each month
  scale_color_manual(values = color_palette) +  # Use a color palette for all months
  labs(
    x = "Day of the Month",
    y = "Count",
    color = "Month"
  ) +
  theme_minimal() +
  theme(
    legend.position = "right"
  )

# Display the plot
flower_date_line_plot
```

@fig-flowerday-2 presents a more detailed view of the flowering dates by month and day. The peak bloom periods are concentrated towards the end of March and throughout April, with the highest counts occurring in late April. This detailed distribution highlights the specific bloom dates, emphasizing how the bulk of flowering happens within a narrow time frame in early spring, largely driven by favorable temperature conditions.

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.pos="H"}
#| label: fig-flowerday-3
#| fig-cap: "Temperature Distribution by Date"

#### Plot Distribution of Temperature by Date ####

# Extract date and temperature from the dataset
temperature_distribution <- sakura_modern_cleaned %>%
  select(flower_date, mean_temp_month) %>%
  mutate(
    day_of_year = yday(flower_date),  # Convert date to day of the year
    month = month(flower_date, label = TRUE, abbr = TRUE)  # Extract month for potential color coding
  )

# Create the plot
temperature_plot <- ggplot(temperature_distribution, aes(x = day_of_year, y = mean_temp_month)) +
  geom_point(alpha = 0.5, color = "steelblue") +  # Scatter plot of temperature
  geom_smooth(method = "loess", color = "darkorange", size = 1, se = FALSE) +  # Smoothed line for trend
  labs(
    x = "Day of the Year",
    y = "Temperature (°C)"
  ) +
  scale_x_continuous(
    breaks = c(1, 32, 60, 91, 121, 152),  # Manually specify breaks for Jan, Feb, Mar, Apr, May
    labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun")  # Correct labels
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 12),
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12)
  )

# Display the plot
print(temperature_plot)

```

@fig-flowerday-3 shows the relationship between day of the year (DOY) for blooming and temperature. The trend line suggests a potential linear relationship, where later dates are associated with higher temperatures. This trend indicates that as the year progresses, temperatures rise, which influences the timing of cherry blossom blooms. The scatter plot further emphasizes the variation in temperatures experienced during different bloom periods, illustrating how the blooming date affects the observed temperature.

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.pos="H"}
#| label: fig-flowerday-4
#| fig-cap: "Average Flowering Dates Across Japan"

#### Load and Prepare Data ####
# Calculate mean flower_doy for each location
sakura_summary <- sakura_modern_cleaned %>%
  group_by(latitude, longitude) %>%
  summarise(mean_flower_doy = mean(flower_doy, na.rm = TRUE)) %>%
  ungroup()

#### Load Japan's Basemap ####
japan_map <- ne_countries(scale = "medium", country = "Japan", returnclass = "sf")

#### Create the Map ####
gis_plot <- ggplot(data = japan_map) +
  geom_sf(fill = "white", color = "black") +  # Base map of Japan
  geom_point(
    data = sakura_summary,
    aes(x = longitude, y = latitude, color = mean_flower_doy),
    size = 3, alpha = 0.7
  ) +
  scale_color_viridis_c(
    option = "D",  # Viridis color scale for better visualization
    name = "Mean Flower DOY"
  ) +
  theme_minimal() +
  labs(
    subtitle = "Darker points indicate later flowering dates",
    x = "Longitude",
    y = "Latitude"
  ) +
  theme(legend.position = "right")

# Display the plot
print(gis_plot)
```

@fig-flowerday-4 provides a geographic overview of average flowering dates across Japan. The color gradient represents the mean flowering DOY, with darker points indicating later flowering dates. This map illustrates how flowering dates vary geographically, with earlier blooms occurring in warmer, southern regions and later blooms in cooler, northern regions. This geographic variation in bloom dates underscores the influence of temperature gradients across Japan.

In summary, the timing of cherry blossom blooms significantly affects temperature patterns. The majority of cherry blossoms bloom in March and April, which in turn correlates with observed temperature trends. Geographic variation also plays a significant role, with southern regions experiencing earlier blooms compared to northern regions. These insights are essential for understanding the temporal dynamics of sakura florescence and for enhancing the accuracy of predictive models that aim to understand temperature dynamics under varying bloom conditions. These insights are essential for understanding the temporal dynamics of sakura florescence and for enhancing the accuracy of predictive models that aim to forecast bloom timing under varying climatic conditions.

## Excluded Variables

### Year

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.pos="H"}
#| label: fig-year
#| fig-cap: "Temperature Trend Over Years"

#### Prepare Data ####
# sakura_historical_cleaned is already loaded
line_data <- sakura_historical_cleaned %>%
  filter(!is.na(temp), !is.na(year))  # Ensure no missing values

# Calculate mean temperature to use as a horizontal reference line
mean_temp <- mean(line_data$temp, na.rm = TRUE)

#### Create Line Plot ####
line_plot <- ggplot(line_data, aes(x = year, y = temp)) +
  geom_line(color = "steelblue", size = 1) +  # Line graph with color and size
  geom_hline(
    yintercept = mean_temp, 
    color = "darkorange", 
    linetype = "dashed", 
    size = 1
  ) +  # Add horizontal dashed line for reference
  labs(
    x = "Year",
    y = "Temperature (°C)"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 12),
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12)
  )

# Display the plot
print(line_plot)

```

@fig-year depicts the temperature trend over several centuries, highlighting the distribution of temperatures across different years. The orange dashed line represents the average temperature, providing a baseline for comparison. The plot shows considerable variation in temperatures from year to year, with no consistent linear trend. This lack of a clear relationship between temperature and year suggests that, over the long term, temperature fluctuations have been influenced by a complex interplay of factors beyond a simple time progression.

While there are periods of both high and low temperature variability, the data does not support a strong linear trend correlating temperature with time. This finding emphasizes that year-to-year temperature changes are not solely driven by temporal progression, but rather are subject to regional and potentially global climatic influences. Understanding these temperature trends is crucial for predicting future cherry blossom bloom timings, as temperature is a key determinant in bloom dynamics.

### Longitude

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.pos="H"}
#| label: fig-longitude-1
#| fig-cap: "Longitude Distribution"

#### Improved Plot Latitude and Longitude Distributions ####
# Plot longitude distribution with bins
longitude_plot <- ggplot(sakura_modern_cleaned, aes(x = longitude)) +
  geom_histogram(
    bins = 20,                    # Adjust the number of bins for better grouping
    fill = "darkorange", 
    alpha = 0.8
  ) +
  labs(
    x = "Longitude",
    y = "Count"
  ) +
  theme_minimal()

# Display the plots
longitude_plot

```

@fig-longitude-1 shows the distribution of longitudes for the locations included in the dataset. The majority of the observations are concentrated between 135°E and 141°E, representing regions across central and eastern Japan. This distribution highlights the geographic focus of our study, which includes areas where cherry blossoms are commonly observed and where temperature data is readily available.

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.pos="H"}
#| label: fig-longitude-2
#| fig-cap: "Temperature VS Longitude"

#### Prepare Longitude Data ####
longitude_temp_data <- sakura_modern_cleaned %>%
  filter(!is.na(longitude), !is.na(mean_temp_month)) %>%
  group_by(longitude) %>%
  summarise(mean_temp = mean(mean_temp_month, na.rm = TRUE)) %>%
  ungroup()

#### Fit a Linear Model ####
linear_model <- lm(mean_temp ~ longitude, data = longitude_temp_data)

#### Create Plot ####
longitude_temp_plot <- ggplot(longitude_temp_data, aes(x = longitude, y = mean_temp)) +
  geom_line(color = "darkorange", size = 1) +  # Line graph with custom color and size
  geom_abline(
    slope = coef(linear_model)["longitude"],
    intercept = coef(linear_model)["(Intercept)"],
    color = "steelblue",
    linetype = "dashed",
    size = 1
  ) +  # Add a dashed line for the linear relationship
  labs(
    x = "Longitude",
    y = "Mean Temperature (°C)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 12),
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12)
  )

# Display the plot
print(longitude_temp_plot)

```

@fig-longitude-2 illustrates the relationship between longitude and mean temperature. The dashed blue line represents the linear trend, indicating a positive correlation between longitude and temperature. However, compared to latitude, this relationship is weaker and exhibits more variability. The positive trend suggests that regions further east tend to have slightly higher average temperatures, though the variability indicates that other geographic and climatic factors are also at play.

In summary, while longitude does show some relationship with temperature, it is not as strong as the correlation observed with latitude. This weaker relationship and greater variability led us to exclude longitude as a predictor in Model 2. The rationale for this decision will be discussed in detail in @sec-model, where we explain the selection of the most relevant variables for predicting temperature and bloom dynamics.

# Model {#sec-model}

## Alternative Model

```{=tex}
\begin{equation}
\text{Alternative Model: } \text{mean\_temp\_month} = \beta_1 \cdot \text{latitude} + \beta_2 \cdot \text{longitude} + \beta_0
\label{eq:alternative_model}
\end{equation}

\noindent
where:
\begin{itemize}
    \item $\text{latitude} \in [20, 50]$, representing geographical latitude in degrees (°N).
    \item $\text{longitude} \in [120, 150]$, representing geographical longitude in degrees (°E).
    \item $\beta_1$ and $\beta_2$ are coefficients for $\text{latitude}$ and $\text{longitude}$, respectively.
    \item $\beta_0$ is the intercept term.
    \item Both $\text{latitude}$ and $\text{longitude}$ are numerical variables.
\end{itemize}
```

As shown in equation~\ref{eq:alternative_model},the alternative model was developed to investigate whether latitude and longitude can simultaneously maintain a linear relationship with mean temperature. Specifically, this model aims to evaluate if both geographic coordinates contribute meaningfully to temperature estimation, thereby affecting the prediction accuracy for bloom timing.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#| label: tbl-am
#| tbl-cap: "Alternative Model Summary"

# Create a data frame for the coefficients and p-values
summary_table <- data.frame(
  Variable = c("latitude", "longitude"),
  Estimate = c(0.778101, -0.003319),
  P_Value = c("<2e-16", "0.806")
)

# Add R-squared information to the table
r_squared <- 0.6722
adjusted_r_squared <- 0.6721

# Combine R-squared information into the table
model_info <- data.frame(
  Variable = c("R-squared", "Adjusted R-squared"),
  Estimate = c(r_squared, adjusted_r_squared),
  P_Value = c("", "") # No p-value for R-squared metrics
)

# Combine both tables
final_table <- rbind(summary_table, model_info)

# Beautify the final_table using kable
kable(
  final_table,
  format = "latex",             # Output in LaTeX format
  booktabs = TRUE,              # Use booktabs style for clean table
  align = "c",                  # Center align columns
  col.names = c("Variable", "Estimate", "P-Value") # Column names
) %>%
  kable_styling(
    latex_options = c("striped", "hold_position"), # Add striped rows and fixed position
    full_width = FALSE                             # Keep the table compact
  ) %>%
  add_header_above(c(" " = 1, "Model Summary" = 2)) # Add a header spanning the last two columns
```

@tbl-am provides the summary of the alternative model, showing the estimates and p-values for both latitude and longitude. The results indicate that while latitude has a strong and statistically significant relationship with temperature, longitude does not contribute significantly to the model (p-value of 0.806). This suggests that latitude is the more important predictor of mean temperature, which aligns with our earlier analysis showing the stronger linear relationship between latitude and temperature compared to longitude.

The findings from this alternative model indicate that incorporating longitude as a predictor does not substantially improve temperature estimation. Consequently, we excluded longitude in our final temperature estimation model (Model 2), focusing solely on latitude, month, and date as predictors. This decision is based on both statistical performance and model simplicity, aiming to create a model that is interpretable without sacrificing predictive accuracy.

## Model 1

Model 1,as shown in equation~\ref{eq:model_1}, is designed to predict the duration of cherry blossom blooms, represented as flower_doy (day of year), based on the temperature variable (temp). This model aims to understand how temperature directly influences the timing and duration of sakura blooms.

```{=tex}
\begin{equation}
\text{Model 1: } \text{flower\_doy} = \beta_1 \cdot \text{temp} + \beta_0
\label{eq:model_1}
\end{equation}

\noindent
where:
\begin{itemize}
    \item $\text{temp} \in [1, 12]$, representing the temperature range in degrees Celsius (°C).
    \item $\beta_1$ is the coefficient of the variable $\text{temp}$, and $\beta_0$ is the intercept term.
    \item The variable $\text{temp}$ is a numerical variable.
\end{itemize}
```

The primary goal of Model 1 is to determine the effect of temperature on the flowering date of cherry blossoms. By focusing on the temperature variable alone, this model isolates the direct impact of climatic conditions on bloom timing. This is particularly relevant in the context of climate change, as increasing temperatures may significantly alter bloom patterns. The simplicity of Model 1 also ensures that the results are interpretable, making it easier to draw conclusions regarding the role of temperature in bloom duration.

Model 1 uses temperature as the only predictor, making it a straightforward linear model. This approach helps establish a foundational understanding of the direct relationship between temperature and bloom timing before introducing additional complexities. The linear form of the model allows us to quantify the influence of temperature on bloom day and evaluate whether temperature changes can be directly linked to shifts in flowering periods. This model provides a baseline against which more complex models can be compared, offering insights into the primary climatic driver of cherry blossom florescence.

@tbl-model1 summarizes the results of Model 1, including residuals, coefficients, and overall model performance metrics. A detailed breakdown of model diagnostics and performance metrics is available in @sec-appendix for further reference.

## Model 2

Model 2,as shown in equation~\ref{eq:model_2}, is designed to estimate the mean monthly temperature based on three predictors: day of the month (day), latitude, and month. This model aims to establish how temperature can be estimated as a function of both spatial and temporal variables, which then informs the bloom prediction model.

```{=tex}
\begin{equation}
\text{Model 2: } \text{mean\_temp\_month} = \beta_1 \cdot \text{day} + \beta_2 \cdot \text{latitude} + \beta_3 \cdot \text{month} + \beta_0
\label{eq:model_2}
\end{equation}

\noindent
where:
\begin{itemize}
    \item $\text{latitude} \in [20, 50]$, representing geographical latitude in degrees (°N).
    \item $\text{day} \in [1, 31]$, representing the day of the month.
    \item $\text{month}$ is a categorical variable representing months (January to May).
    \item $\beta_1$, $\beta_2$, $\beta_3$ are coefficients of the linear model, and $\beta_0$ is the intercept term.
    \item All variables are numerical variables, except for $\text{month}$, which is a categorical variable.
\end{itemize}
```

Model 2 provides a framework to estimate temperature as a function of time and geographical location. By incorporating both latitude and day as numerical variables, and treating month as a categorical variable, the model captures both spatial and seasonal variations in temperature. This is critical for understanding the temperature profile across different regions of Japan and during different periods of the year.

The inclusion of latitude allows Model 2 to account for the spatial gradient in temperature, which is crucial given Japan's north-south orientation and diverse climatic conditions. Additionally, the day variable helps in capturing finer temporal variations within each month. Month, as a categorical variable, enables the model to distinguish between the broad seasonal effects on temperature.

@tbl-model2 provides a summary of Model 2, including residuals, coefficients, and overall model performance metrics. The detailed analysis and diagnostics of Model 2 are provided in  @sec-appendix, where we evaluate its accuracy and predictive capability in estimating temperature based on the selected predictors.

# Result {#sec-result}

## Model Result & Validation

To validate the performance of our models, we split the cleaned data into training and testing sets with a 70-30 ratio. The training set was used to train the models, while the testing set was used for validation. The following results demonstrate the predictive accuracy of the models when applied to the testing set.The results of our modeling efforts are summarized in @fig-result-1 and @fig-result-2, which illustrate the predictive performance of Model 1 and Model 2, respectively.

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.pos="H"}
#| label: fig-result-1
#| fig-cap: "Model 1: Actual vs Predicted Sakura Florescence"

#### Preprocess Test Data for Model 2 ####
# Ensure necessary columns are present for Model 2
latitude_time_test_data <- latitude_time_test_data %>%
  mutate(
    month = factor(month(flower_date)),  # Convert month to a factor (categorical variable)
    day = day(flower_date)              # Extract the day from the flower_date
  )

#### Predictions ####
# Predict using Model 1
temp_test_data <- temp_test_data %>%
  mutate(predicted_flower_doy = predict(model1, newdata = temp_test_data))

#### Plot Results ####
# Plot for Model 1: Actual vs Predicted Flower DOY
plot_model1 <- ggplot(temp_test_data, aes(x = flower_doy, y = predicted_flower_doy)) +
  geom_point(color = "steelblue", alpha = 0.7) +
  geom_abline(slope = 1, intercept = 0, color = "darkorange", linetype = "dashed") +
  labs(
    x = "Actual Flower DOY",
    y = "Predicted Flower DOY"
  ) +
  theme_minimal()

# Display the plots
print(plot_model1)
```

@fig-result-1 depicts the actual versus predicted flowering day of the year (DOY) for cherry blossoms using Model 1. The points are closely aligned along the dashed orange line, indicating that the model predictions are highly accurate. The strong linear relationship observed suggests that temperature is an effective predictor of bloom timing, validating our choice of using temperature as the sole predictor in Model 1.The coefficient of variables in model 1 and corresponding p-value is listed below in @tbl-model1-coefficients. Full summary table is in @sec-appendix.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#| label: tbl-model1-coefficients
#| tbl-cap: "Coefficients for Model 1"

# Create a data frame for coefficients
coefficients_table <- data.frame(
  Variable = c("(Intercept)", "temp"),
  Estimate = c(126.202188, -3.509279),
  Std_Error = c(0.034396, 0.005355),
  T_value = c(3669.1, -655.3),
  P_Value = c("<2e-16", "<2e-16")
)

# Coefficients Table for Model 1
kable(
  coefficients_table, 
  format = "latex", 
  booktabs = TRUE, 
  align = "c",
  col.names = c("Variable", "Estimate", "Std. Error", "t Value", "P-Value")
) %>%
  kable_styling(latex_options = c("striped", "hold_position")) %>%
  add_header_above(c("Coefficients for Model 1" = 5)) 
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.pos="H"}
#| label: fig-result-2
#| fig-cap: "Model 2: Actual vs Predicted Temperature"

# Predict using Model 2
latitude_time_test_data <- latitude_time_test_data %>%
  mutate(predicted_mean_temp_month = predict(model2, newdata = latitude_time_test_data))

# Plot for Model 2: Actual vs Predicted Mean Temperature
plot_model2 <- ggplot(latitude_time_test_data, aes(x = mean_temp_month, y = predicted_mean_temp_month)) +
  geom_point(color = "steelblue", alpha = 0.7) +
  geom_abline(slope = 1, intercept = 0, color = "darkorange", linetype = "dashed") +
  labs(
    x = "Actual Mean Temperature (°C)",
    y = "Predicted Mean Temperature (°C)"
  ) +
  theme_minimal()

print(plot_model2)
```
@fig-result-2 illustrates the actual versus predicted mean temperature values for Model 2.While there is more variability compared to Model 1，the overall trend aligns well with the dashed orange line, indicating reasonable predictive performance. This can be attributed to the increased complexity of the model and the influence of multiple interacting factors, such as day, latitude, and month, which contribute to temperature variations, which is discussed in @sec-discussion. Nonetheless, the model performs well in capturing the overall pattern of temperature change.The coefficient of variables in model 2 and corresponding p-value is listed below in @tbl-model2-coefficients. Full summary table is in @sec-appendix.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#| label: tbl-model2-coefficients
#| tbl-cap: "Coefficients for Model 2"

# Load necessary library
library(knitr)
library(kableExtra)

# Create a data frame for coefficients
coefficients_table <- data.frame(
  Variable = c("(Intercept)", "day", "latitude", "month2", "month3", "month4", "month5"),
  Estimate = c(-9.000439, -0.049223, 0.189083, -0.617982, 2.869383, 6.966108, 10.047363),
  Std_Error = c(0.289464, 0.003291, 0.012089, 0.219831, 0.114364, 0.164984, 0.237453),
  T_value = c(-31.093, -14.956, 15.637, -2.811, 25.088, 42.223, 42.313),
  P_Value = c("<2e-16", "<2e-16", "<2e-16", "0.00496", "<2e-16", "<2e-16", "<2e-16")
)

# Coefficients Table for Model 2
kable(
  coefficients_table, 
  format = "latex", 
  booktabs = TRUE, 
  align = "c",
  col.names = c("Variable", "Estimate", "Std. Error", "t Value", "P-Value")
) %>%
  kable_styling(latex_options = c("striped", "hold_position")) %>%
  add_header_above(c("Coefficients for Model 2" = 5)) 
```

In conclusion, both models show strong predictive capabilities, with Model 1 excelling in predicting bloom timing based on temperature and Model 2 effectively estimating temperature based on geographic and temporal factors. The results demonstrate the utility of the dual-model approach in enhancing our understanding of the factors driving cherry blossom bloom duration, particularly in the context of climate variability and change.

## Example of Prediction

To provide a practical example of how our models perform, we selected a specific data point from the raw dataset representing the city of Esashi in 1969. The data for this location included all relevant attributes needed for the models, such as latitude, day, month, and mean temperature. Using these values, we predicted the mean temperature and flowering day of the year (DOY), then compared the predicted values to the actual historical data.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#| label: tbl-eop
#| tbl-cap: "Prediction Compared to Actual Data"

#### Define Real Data Point ####
real_data <- tibble(
  year = 1969,
  flower_date = as.Date("1969-04-29"),
  flower_doy = 119,
  station_name = "Esashi",
  mean_temp_month = 3.9,
  latitude = 40.82,
  longitude = 140.77
)

#### Prepare Data for Model 2 ####
# Add derived variables: month and day
real_data <- real_data %>%
  mutate(
    month = factor(month(flower_date)),  # Convert month to categorical variable
    day = day(flower_date)              # Extract day of the month
  )

#### Predict Temperature using Model 2 ####
predicted_mean_temp <- predict(model2, newdata = real_data)

#### Prepare Data for Model 1 ####
# Create a new data frame for Model 1 with the predicted temperature
real_data_model1 <- real_data %>%
  mutate(temp = predicted_mean_temp)  # Add predicted temperature as 'temp'

#### Predict Flower DOY using Model 1 ####
predicted_flower_doy <- predict(model1, newdata = real_data_model1)

#### Combine Results ####
comparison <- tibble(
  Metric = c("Mean Temperature (°C)", "Flower DOY"),
  Actual = c(real_data$mean_temp_month, real_data$flower_doy),
  Predicted = c(predicted_mean_temp, predicted_flower_doy)
)

#### Beautify the comparison table using kable ####
comparison %>%
  kable(
    format = "latex",            # Output in LaTeX format
    booktabs = TRUE,             # Use booktabs style for clean table
    align = "c",                 # Center align columns
    col.names = c("Metric", "Actual", "Predicted")  # Column names
  ) %>%
  kable_styling(
    latex_options = c("striped", "hold_position"),  # Add striped rows and fixed position
    full_width = FALSE                              # Keep the table compact
  ) %>%
  add_header_above(c(" " = 1, "Prediction Summary" = 2))  # Add a header spanning last two columns

```

@tbl-eop presents the actual versus predicted values for mean temperature and flowering DOY for the selected data point. The results show that the predicted mean temperature is 4.25°C, compared to the actual value of 3.9°C. For flowering DOY, the model predicted a value of 111.27, while the actual recorded value was 119. This example demonstrates that our models are highly accurate in predicting both temperature and bloom timing, effectively capturing the relationship between environmental factors and cherry blossom florescence.This prediction example highlights the practical application of the models developed in this study and their potential to provide meaningful insights into the timing and conditions of cherry blossom blooms in Japan.

# Discussion {#sec-discussion}

## Limitation

Despite the promising results of our models, several limitations must be acknowledged. First, the models rely heavily on the availability and accuracy of historical and modern phenological data. Data gaps, inconsistencies, or inaccuracies could affect model performance, particularly for predictions in regions or timeframes where data is sparse. Second, while temperature was found to be a strong predictor of bloom timing, it is not the only environmental factor influencing cherry blossom florescence. Other climatic and ecological variables, such as precipitation, soil moisture, and light exposure, could also play significant roles but were not included in our models due to data limitations.

Additionally, Model 2's prediction results showed some deviations, as seen in @fig-result-2. This suggests that there may be other factors, such as regional or geographic influences, impacting our temperature predictions. These factors could include microclimates, urban heat islands, or variations in local geography, which were not captured in our model. Addressing these additional influences could enhance the model's accuracy in future iterations.

## Interpretation and Future Research Directions

The findings from this study provide valuable insights into the factors that influence cherry blossom bloom timing in Japan. One key takeaway is the strong relationship between temperature and bloom duration, confirming the critical role of climatic conditions in determining the timing of sakura florescence. This information is particularly relevant in the context of climate change, as rising temperatures could lead to earlier and potentially shorter bloom periods, impacting both cultural and ecological aspects of cherry blossom events.

Another important finding is the relative importance of geographic factors, such as latitude, in determining local temperatures. Our results suggest that latitude plays a more significant role than longitude in temperature estimation, leading us to exclude longitude from the final model. This decision reflects a balance between model complexity and predictive accuracy, ensuring that the model remains interpretable while retaining its predictive power.

Future research should explore additional environmental variables that may affect bloom timing, such as soil moisture, precipitation, and light availability. Incorporating these factors could improve model accuracy and provide a more comprehensive understanding of the factors driving sakura florescence. Furthermore, future studies could consider non-linear models or machine learning approaches to capture complex interactions between variables, potentially enhancing predictive capabilities.

Finally, it would be beneficial to extend the scope of the analysis beyond Japan to examine cherry blossom bloom patterns in other regions, such as South Korea or China. By comparing bloom dynamics across different climates and geographies, we could gain a deeper understanding of the global factors affecting cherry blossom phenology and assess the broader implications of climate change on these iconic events.

# Appendix {#sec-appendix}

## Survey Methodology Overview

In this section, we provide a comprehensive overview of the survey methodologies relevant to this study, focusing on their design, sampling strategies, and implications for data quality. Surveys serve as a critical tool for capturing public sentiment and translating it into actionable insights, especially in environmental research where individual perceptions and regional trends play a crucial role.

Our primary data collection involved historical and modern records of cherry blossom blooming and temperature data. To complement these observations, an ideal survey could be implemented to gauge local experiences of sakura blooming and capture potential environmental influences not represented in raw data. This survey would be designed using a combination of stratified random sampling and purposive sampling, ensuring representation across various geographic regions and demographics within Japan.

The survey would be stratified by latitude and elevation bands, recognizing that blooming times differ significantly across these dimensions. This approach would ensure our sample includes participants from urban, rural, coastal, and inland areas, providing a richer context for the impact of climatic changes on cherry blossom timing. Sampling would also consider socio-demographic factors such as age and occupation, as these factors may influence awareness and historical knowledge of bloom events. Data collection methods would include both online surveys and in-person interviews to maximize reach and accessibility.

Moreover, measurement tools within the survey would follow well-established practices. Questions would address individuals' observations on bloom timing, any perceived shifts over years, and associated environmental conditions. The survey would utilize a Likert scale to capture perceptions of climatic changes, and open-ended questions to gather qualitative insights into local conditions and environmental anecdotes that may not be captured by weather stations.

## Idealized Survey

To supplement the data-driven models presented in this paper, we propose an idealized survey that could provide more in-depth insights into the local experiences of climate effects on cherry blossom blooming. If we had an extended budget of $100,000, this survey would be expanded to cover additional aspects not readily accessible through observational or temperature datasets alone.

The idealized survey would be conducted over a full blooming season, collecting data from diverse regions across Japan. The goal would be to capture spatial and temporal variability in bloom timing and link it to local microclimatic conditions, participant observations, and other environmental factors such as soil moisture and local flora interactions. We would employ a mixed-methods approach, combining both quantitative (closed questions) and qualitative (interviews, focus groups) data.

Sampling would be implemented through a mixture of probability and non-probability approaches to achieve both representativeness and depth. Stratified random sampling would be used to ensure geographic diversity, while snowball sampling would help gather data from specific cultural or expert groups (e.g., horticulturalists, elders who possess traditional knowledge about historical blooming patterns). Each participant would provide data on observed bloom dates, perceived changes in bloom patterns over time, and any unusual climate events that they believe might have influenced sakura blooming.

Survey questions would be developed with guidance from both climate scientists and cultural experts to ensure that they adequately capture the climatic, geographic, and cultural contexts of sakura blooming. Likert scales would be used to gauge perceptions of change in bloom patterns and the impact of factors such as temperature and rainfall. In addition, open-ended questions would allow respondents to describe any unusual events or trends they have noticed over time.

The survey results would be analyzed using statistical techniques, including logistic regression to identify factors associated with changes in bloom timing, and thematic analysis for qualitative responses. The insights gained would be used to refine our predictive models by incorporating more nuanced, location-specific environmental factors that are not currently represented in our dataset.

The implementation of this idealized survey could significantly enhance the robustness of our models, offering deeper insights into the role of microclimates and local environmental influences on cherry blossom phenology. Moreover, it would provide a valuable opportunity to include public perception in our analysis, offering a human dimension to the understanding of climatic impacts on a culturally significant event.

## Survey Design

To provide a comprehensive understanding of local environmental factors impacting cherry blossom blooming, we propose a survey targeted at residents across Japan. This survey aims to gather data regarding personal observations of bloom dates, temperature changes, and any noted environmental anomalies. Below is a suggested questionnaire to gather insights from participants.

1. **What is your age group?**
    - Under 18
    - 18-34
    - 35-49
    - 50-64
    - 65 and above

2. **What is your gender?**
    - Male
    - Female
    - Prefer not to say

3. **In which region of Japan do you currently reside? (Select one)**
    - Hokkaido
    - Tohoku
    - Kanto
    - Other: ______ (PleaseTypeHere)

4. **How long have you been observing cherry blossom bloom in your area?**
    - Less than 1 year
    - 1-5 years
    - More than 5 years

5. **Have you noticed any changes in the blooming dates over the past years?**
    - Yes, blooming is earlier
    - Yes, blooming is later
    - No significant change
    - Not sure

6. **Which environmental factors do you think have the most impact on bloom timing? (Select all that apply)**
    - Temperature changes
    - Rainfall levels
    - Urban development
    - Other (please specify)

7. **On average, when do cherry blossoms begin to bloom in your area?**
    - Before March
    - March (early)
    - March (mid)
    - March (late)
    - April (early)
    - April (mid)
    - April (late)
    - After April

8. **How would you describe the temperature during the blooming period in the past few years?**
    - Much warmer
    - Slightly warmer
    - About the same
    - Slightly cooler
    - Much cooler

9. **Have there been any unusual climate events (e.g., heavy snowfall, prolonged warm spells) in recent years that might have affected blooming?**
    - Yes (please specify)
    - No
    - Not sure

10. **Do you think climate change is influencing the blooming pattern of cherry blossoms in your region?**
    - Strongly agree
    - Agree
    - Neutral
    - Disagree
    - Strongly disagree

11. **Are there other plants in your region that show similar blooming patterns to cherry blossoms?**
    - Yes (please specify)
    - No
    - Not sure

## Model Details

This section provides full summary table for model 1 and model 2.

\clearpage

```{r, echo=FALSE, message=FALSE, warning=FALSE, tbl.pos="H"}
#| label: tbl-model1
#| tbl-cap: "Model 1 Summary Tables"

# Create a data frame for residuals
residuals_table <- data.frame(
  Statistic = c("Min", "1Q", "Median", "3Q", "Max"),
  Value = c(-0.25208, -0.14651, -0.04095, 0.04606, 1.30709)
)

# Create a data frame for coefficients
coefficients_table <- data.frame(
  Variable = c("(Intercept)", "temp"),
  Estimate = c(126.202188, -3.509279),
  Std_Error = c(0.034396, 0.005355),
  T_value = c(3669.1, -655.3),
  P_Value = c("<2e-16", "<2e-16")
)

# Create a data frame for model summary
model_summary_table <- data.frame(
  Metric = c(
    "Residual Standard Error", 
    "Multiple R-squared", 
    "Adjusted R-squared", 
    "F-statistic", 
    "P-value"
  ),
  Value = c(
    "0.2337 on 572 degrees of freedom",
    "0.9987",
    "0.9987",
    "4.294e+05 on 1 and 572 DF",
    "<2e-16"
  )
)

# Residuals Table for Model 1
kable(
  residuals_table, 
  format = "latex", 
  booktabs = TRUE, 
  align = "c"
) %>%
  kable_styling(latex_options = c("striped", "hold_position")) %>%
  add_header_above(c("Residuals for Model 1" = 2))

# Coefficients Table for Model 1
kable(
  coefficients_table, 
  format = "latex", 
  booktabs = TRUE, 
  align = "c",
  col.names = c("Variable", "Estimate", "Std. Error", "t Value", "P-Value")
) %>%
  kable_styling(latex_options = c("striped", "hold_position")) %>%
  add_header_above(c("Coefficients for Model 1" = 5))

# Model Summary Table for Model 1
kable(
  model_summary_table, 
  format = "latex", 
  booktabs = TRUE, 
  align = "c",
  col.names = c("Metric", "Value")
) %>%
  kable_styling(latex_options = c("striped", "hold_position")) %>%
  add_header_above(c("Model Summary for Model 1" = 2))

```

\clearpage

```{r, echo=FALSE, message=FALSE, warning=FALSE, tbl.pos="H"}
#| label: tbl-model2
#| tbl-cap: "Model 2 Summary Tables"

# Create a data frame for residuals
residuals_table <- data.frame(
  Statistic = c("Min", "1Q", "Median", "3Q", "Max"),
  Value = c(-4.2185, -0.6993, -0.0729, 0.6641, 4.1114)
)

# Create a data frame for coefficients
coefficients_table <- data.frame(
  Variable = c("(Intercept)", "day", "latitude", "month2", "month3", "month4", "month5"),
  Estimate = c(-9.000439, -0.049223, 0.189083, -0.617982, 2.869383, 6.966108, 10.047363),
  Std_Error = c(0.289464, 0.003291, 0.012089, 0.219831, 0.114364, 0.164984, 0.237453),
  T_value = c(-31.093, -14.956, 15.637, -2.811, 25.088, 42.223, 42.313),
  P_Value = c("<2e-16", "<2e-16", "<2e-16", "0.00496", "<2e-16", "<2e-16", "<2e-16")
)

# Create a data frame for model summary
model_summary_table <- data.frame(
  Metric = c(
    "Residual Standard Error", 
    "Multiple R-squared", 
    "Adjusted R-squared", 
    "F-statistic", 
    "P-value"
  ),
  Value = c(
    "1.116 on 4256 degrees of freedom",
    "0.9255",
    "0.9254",
    "8814 on 6 and 4256 DF",
    "<2e-16"
  )
)

# Residuals Table
kable(
  residuals_table, 
  format = "latex", 
  booktabs = TRUE, 
  align = "c"
) %>%
  kable_styling(latex_options = c("striped", "hold_position")) %>%
  add_header_above(c("Residuals for Model 2" = 2))

# Coefficients Table
kable(
  coefficients_table, 
  format = "latex", 
  booktabs = TRUE, 
  align = "c",
  col.names = c("Variable", "Estimate", "Std. Error", "t Value", "P-Value")
) %>%
  kable_styling(latex_options = c("striped", "hold_position")) %>%
  add_header_above(c("Coefficients for Model 2" = 5))

# Model Summary Table
kable(
  model_summary_table, 
  format = "latex", 
  booktabs = TRUE, 
  align = "c",
  col.names = c("Metric", "Value")
) %>%
  kable_styling(latex_options = c("striped", "hold_position")) %>%
  add_header_above(c("Model Summary for Model 2" = 2))

```

\clearpage

# References